---
title: "Temporary File for Two-box Bayesian Model"
author: "Anna Vincent"
date: "2025-01-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Set working directory and load packages, message=FALSE}

rm(list=ls())
setwd("/Users/annavincent/Documents/Bayesian Collab/LeefBayes")
getwd()

#devtools::install_github("wilkelab/ungeviz")

library(rstan)
options(mc.cores = parallel::detectCores()) #run stan in parallel
#library(shinystan)
library(tidyverse)
library(plyr)
library(tidyr)
library(tibble)
library(lubridate)
library(broom)
library(dplyr)
library(data.table)
library(grid)
library(ungeviz)

```

## The Original Model
Vincent et al. (2025), JGR: Biogeosciences (DOI: 10.1029/2024JG008259)

We took the mean of the the 4 stream types (labelled as 1-4) in a 3-level (hierarchical) model. Error enters at 1) the within-release level, 2) the among-release level for any one stream (i.e. variation in removal rate $k$), and 3) the variation among the 4 streams... Put simply, error enters the dataset/model at 3 different levels: 1) within a given release, 2) within releases for a given stream, and 3) across streams.

The first level was
 
$$A_{i,j} = A_0 \times e^{-k_j x} + \varepsilon_A$$
$$\varepsilon_A \sim N(0,\sigma_A)$$
$$N_{i,j} = N_0 \times e^{-k_j x} + \varepsilon_N$$
$$\varepsilon_N \sim N(0,\sigma_N)$$
This model estimated $k_j$, which is the per m decline in either NH4 ($A$) or NO3 ($N$) as a function of distance ($x$) and is specific for a release ($j$). In these equations, $A_{i,j}$ and $N_{i,j}$ represent [NH4] or [NO3], respectively, at a given station ($i$) during a given release ($j$). It is based on [NH4] or [NO3] at the top station ($A_0$ or $N_0$; which we used to normalize the release data for subsequent stations). We assumed an exponential decline ($e$) and normally distributed errors (equations for $\varepsilon_A$ or $\varepsilon_N$).

We use $k$ estimates to calculate uptake velocity $V_f$ of either NH4 or NO3

The next level estimated variation in $V_{fj}$ within individual streams, which is the update velocity for a given release ($j$).

$$V_{f,j,s}= V_{f,s} + b \times biofilm + c \times shade + d \times biofilm*shade + \varepsilon_v$$
We were interested in uptake patterns across 4 biofilm periods (shaded; unshaded: early biofilm; unshaded: late biofilm; reshaded). In our experiment, dates 1 & 4 (7/1/20, 8/15/20) were shaded and reshaded, respectively; dates 2 & 3 were early and late biofilm, respectively. We used a binary (0,1) system to code this into our dataset, where shaded and late biofilm colonization treatments received a "1" 

$$ \varepsilon_v \sim N(0,\sigma_v)$$
Lastly we model the among stream variation in $V_{fs}$ as
$$V_{f,s}= V_{f, mean} + \varepsilon_s$$
$$ \varepsilon_s \sim N(0,\sigma_s)$$


## Generating a fake dataset

```{r Creation of fake dataset}

Q <- 0.001*86400  #m3/d --> 1 L is 0.001 m3; multiply by 86400 to get from seconds (L/s) to days (m3/d)
w <- 0.5 #stream width (m)

n_stream <- 4
n_release <- 12 #number of releases per stream

#start at the top, work down
meanvf <- log(3) #m/d
sdvf <- 0.1 # stream to stream differences in vf; parameter, not an error.  Note that log std is very different than sd.  It is like fraction erro, and scale independent
vfstream <- rnorm(n_stream,meanvf,sdvf)

sd_release <- 0.2 ##much smaller for logs.  This says "10% error"
vf <- rep(vfstream, each=12) + rnorm(n_stream*n_release, 0, sd_release) #now make vf for each release

##now make ammonium uptake

k <- exp(vf)*w/Q #this is correct. See the difference, you exponentiated w and q too in k<- exp(w/(vf*Q)).  I also screwed the algebra before, so this is better

amm_sd <- 0.05 # wow am I good at measuring NH4
stations <- rep(c(10,20,30,40,50), n_stream*n_release)
krep <- rep(k, each=5)
amm_corr <- exp(-krep*stations) + rnorm (length(stations),0, amm_sd) #simulated [NH4] at each station
#

##now make fake dataframe

fakedat <- data.frame(amm_corr=amm_corr, q=Q, stations=stations, stream=rep(1:4,each=60), release=rep(rep(1:12, each=5),4), group=rep(1:48, each=5))

```
